A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-07-08 text/manuscript.Rmd
add Hanahan and Weinberg reference

Other studies aim at providing a comprehensive view of mutations, gene
expression and genomic signatures, with the express goal of understanding the
common themes of all cancer independent of tissue of origin. The first of these
(the “Pan-Cancer-12” analysis) identified 11 subtypes from 12 cancer types,
using integrative analysis with co-equal weighting of gene expression,
methylation, copy number and proteomics data
[@hoadleyMultiplatformAnalysis2014]. The principal finding was that
tissue-of-origin is the predominant driving factor, though a significant
proportion of tumors ($10\%$) could be reclassified independent of
tissue-of-origin. In a second study involving 33 cancer types and a much greater
number of tumors using a newer algorithm (iCluster)
[@shenIntegrativeClusteringMultiple2009], the authors identify 28 clusters that
could be further related or subdivided into organ specific groups, including
pan-gastrointestinal, pan-gynecological, pan-squamous, pan-gynecological/
squamous and pan-kidney [@hoadleyCellofOriginPatternsDominate2018].

>>>>

Other studies aim at providing a comprehensive view of mutations, gene
expression and genomic signatures, with the express goal of understanding the
common themes of all cancer independent of tissue of origin. Such a view of
cancer as a malady affecting common cellular processes has long been a goal of
the field and is perhaps best characterized in an essay by @weinberg2000 (see also
Discussion). The first of these studies, (the
“Pan-Cancer-12” analysis) identified 11 subtypes from 12 cancer types,
using integrative analysis with co-equal weighting of gene expression,
methylation, copy number and proteomics data
[@hoadleyMultiplatformAnalysis2014]. The principal finding was that
tissue-of-origin is the predominant driving factor, though a significant
proportion of tumors ($10\%$) could be reclassified independent of
tissue-of-origin. In a second study involving 33 cancer types and a much greater
number of tumors using a newer algorithm (iCluster)
[@shenIntegrativeClusteringMultiple2009], the authors identify 28 clusters that
could be further subdivided into organ specific groups, including
pan-gastrointestinal, pan-gynecological, pan-squamous, pan-gynecological/
squamous and pan-kidney [@hoadleyCellofOriginPatternsDominate2018].
